---
title: "Volume Estimates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Volume Estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(picoinjector)
library(dplyr)
library(ggplot2)
library(tidyr)
```

In this vignette, calculations to estimate volumes and quantities relevant for the droplet making and picoinjection workflow are documented.


### Tubing

Volume of a cylinder with radius $r$ and height $h$:

$$V = \pi r^2 h$$

```{r}
getCylinderVolume <- function(radius, height) {
  return(pi * radius^2 * height)
}
```



Tubing options


| type   | inner diameter      | outer diameter      |
|--------|---------------------|---------------------|
| Typgon | 0.508 mm (0.020 in) | 1.524 mm (0.060 in) |
| PTFE   | 0.8 mm              | 1.6 mm              |

```{r}
# everything in m
radiusTygon <- 0.508e-3 / 2
radiusPTFE <- 0.8e-3 / 2
```


Plot the tubing volume as a function of tubing length for lengths up to 1 meter:

```{r}
volumes <- tibble::tibble("length" = seq(1,1000)) %>% 
  dplyr::mutate(length = length * 1e-3) %>% # convert into mm
  dplyr::mutate(v_tygon = getCylinderVolume(radiusTygon, length),
                v_ptfe = getCylinderVolume(radiusPTFE, length)) %>%
  tidyr::pivot_longer(c(v_tygon, v_ptfe), names_to = "type", values_to = "volume")
volumes

ggplot(volumes, aes(x = length, y = volume * 1e9, color = type)) + 
  geom_line() + 
  scale_colour_viridis_d(labels = c("PTFE", "Tygon")) + 
  labs(x = "Tubing length [m]",
       y = "Tubing volume [µL]",
       color = "Tubing type") + 
  theme_pretty()
```



Convert flowrate unit µl/h into tubing cm/h:

```{r}
flowrates <- tibble::tibble("ul_h" = seq(0, 100)) %>% 
  dplyr::mutate("Tygon" = ul_h / (getCylinderVolume(radiusTygon, 0.01) * 1e9), 
                "PTFE" = ul_h / (getCylinderVolume(radiusPTFE, 0.01) * 1e9)) %>% 
  tidyr::pivot_longer(!ul_h, names_to = "tubing_type", values_to = "cm_h")
```


```{r}
ggplot(flowrates, aes(x = ul_h, y = cm_h, color = tubing_type)) + 
  geom_line() + 
  scale_color_viridis_d() + 
  theme_pretty() + 
  labs(x = "Flowrate [ul/h]",
       y = "Tubing-length flowrate [cm/h]",
       color = "Tubing type")
```



### Droplets

How much liquid is contained in a droplet of diameter $d$?

Volume of a sphere:
$$V = \frac{4}{3} \pi r^3$$

```{r}
getSphereVolume <- function(radius) {
  return(4/3 * pi * radius^3)
}
```


Plot droplet volumes as function of diameter:

```{r}
tibble::tibble("diameter" = seq(10, 100) * 1e-6) %>% # 10-200 µm droplets
  dplyr::mutate("volume" = getSphereVolume(diameter/2)) %>% 
  ggplot(aes(x = diameter*1e6, y = volume * 1e15)) + # y-axis in picoliters
  geom_line() +
  labs(x = "Droplet diameter [µm]",
       y = "Droplet volume [pL]") + 
  theme_pretty()

```


For droplet sizes up to 40 µm:

```{r}
last_plot() + 
  coord_cartesian(xlim = c(10, 40), ylim = c(0, 35))
```




How many droplets are there in x mm of Tygon/PTFE tubing?


```{r}
radiusDroplet <- 15e-6 # 30 micron droplets
```

Assuming close-packing of equal spheres: fraction of space occupied by spheres is

$$\frac{\pi}{3 \sqrt{2}} \approx 0.74048$$

Using this packing density, I can calculate the volume occupied by spheres for a given tubing length.
Dividing by the known volume of a single droplet yields the total number of droplets contained in x meters of tubing, assuming close-packing of equal spheres.

```{r}
getNumberOfClosePackedSpheres <- function(radius, volume) {
  volumeSingleSphere <- 4/3 * pi * radius^3
  packingDensity <- pi/3/sqrt(2)
  return(packingDensity * volume / volumeSingleSphere)
}
```


Plot number of close-packed spheres over tubing length:

```{r}
volumes %>% 
  dplyr::mutate(n_droplets = getNumberOfClosePackedSpheres(radiusDroplet, volume)) %>% 
  ggplot(aes(x = length, y = n_droplets*1e-6, color = type)) +
  geom_line() +
  scale_colour_viridis_d(labels = c("PTFE", "Tygon")) + 
  labs(x = "Tubing length [m]",
       y = "Number of close-packed spheres [million]",
       color = "Tubing type") + 
  theme_pretty()
```





