---
title: "Electrical Engineering"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Electrical Engineering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Voltage amplifier output measurement


```{r setup, message=FALSE}
library(picoinjector)
library(readr)
library(dplyr)
library(ggplot2)
```


Read in data:

```{r}
d <- read_csv("data/boexle_sorter_output.csv", show_col_types = FALSE)
d
```



Plot:

```{r}
ggplot(d, aes(x = input_vpp, y = output_vpp, color = factor(channel))) + 
  geom_point(shape = 23) + 
  geom_smooth(method = "lm", formula = "y ~ x") + 
  theme_pretty()
```


It seems like variables frequency and channel do not contribute significantly to the output voltage, therefore I am just fitting a single line through all the data points assuming a linear dependency "output voltage ~ input voltage".


```{r}
p <- ggplot(d, aes(x = input_vpp, y = output_vpp)) + 
  geom_point(shape = 23) + 
  geom_smooth(method = "lm", formula = "y ~ x", fullrange = TRUE, color = "black", linewidth = 1) + 
  xlim(c(0, 12.5)) + 
  labs(title = '"boexle sorter" output voltage as function of input voltage',
       x = "Input voltage peak-to-peak",
       y = "Output voltage peak-to-peak") + 
  theme_pretty()
p
```

Prepare a plot that can be printed on A5:

```{r, fig.width=8.27, fig.height=5.83, eval=FALSE}
p + 
  scale_y_continuous(breaks = seq(0, 700, 100), minor_breaks = seq(0, 700, 25)) + 
  scale_x_continuous(breaks = seq(0, 12, 2), minor_breaks = seq(0, 13), limits = c(0, 13))
```


Calculate linear regression manually:

```{r}
lm(output_vpp ~ input_vpp, d)
```

The amplifier gain is therefore 52x.
