---
title: "FADS Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FADS Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(picoinjector)
library(dplyr)
library(readr)
library(ggplot2)
```


Using the LabVIEW software, a TSV file can be recorded during in experiment, where every row corresponds to a droplet that is recognized by the sorter.
For a droplet flowing across the focus spot to be recognized, the signal captured by the PMT has to exceed a defined threshold in terms of amplitude and duration.
This way, droplets can be distinguished from the background signal.

```{r}
fads <- fads_read_tsv("data/fads/220916/pi30v4_1.txt")
fads
```

This function allows to plot the red against the green signal:

```{r}
fads_plot_red_green(fads)
```


Same plot, but with hexagonal glyphs:

```{r}
fads_plot_red_green_hex(fads)
```


Lets look at the signal width as well:

```{r}
fads_plot_width(fads)
```


Plot width vs. time:

```{r}
fads_plot_width_time(fads)
```


Plot signal vs. time:

```{r}
fads_plot_signal_time(fads)
```


Plot droplet frequency over time:

```{r}
fads_plot_frequency(fads)
```












