---
title: "Hydraulic resistance"
output: rmarkdown::html_vignette
bibliography: "bibliography.bib"
vignette: >
  %\VignetteIndexEntry{Hydraulic resistance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(picoinjector)
library(ggplot2)
```

## Straight circular channel

Hydraulic analogy of Ohm's law:

$$\Delta P = R_h Q$$ with $R_h$ hydraulic resistance.

For a (straight) channel with circular cross-section, the hydraulic resistance is defined as $R_h = \frac{8 \mu L}{\pi r^4}$ with $\mu$ dynamic viscosity, $L$ channel length, and $r$ channel radius [@bruus_theoretical_2008].
For water I am using $\mu = 8.9 \cdot 10^{-4} \; \mathrm{Pa \cdot s}$.



```{r}
muWater <- 8.9e-4
radiusTygon <- 0.508e-3 /2
radiusPTFE <- 0.8e-3 /2

getHydraulicResistance <- function(mu, length, radius) {
  return(8 * mu * length / pi / radius^4)
}
```



Plot hydraulic resistance over tubing length:

```{r}
resistance <- tibble::tibble("length" = seq(1,1000) * 1e-3, # unit: meters
               "tygon" = getHydraulicResistance(muWater, length, radiusTygon),
               "ptfe" = getHydraulicResistance(muWater, length, radiusPTFE)) %>% 
  tidyr::pivot_longer(!length, names_to = "type", values_to = "r_hydraulic")
resistance

ggplot(resistance, aes(x = length*1e2, y = r_hydraulic*1e-11, color = type)) + 
  geom_line() + 
  scale_color_viridis_d(labels = c("PTFE", "Tygon")) + 
  labs(x = "Tubing length [cm]",
       y = "Hydraulic resistance [10¹¹ Pa s m⁻³]",
       color = "Tubing type") + 
  theme_pretty()
```





Plot flow rate resulting from pressure drop of 1 Pa as function of tubing length:

```{r}
resistance %>% 
  dplyr::mutate("q" = 1 / r_hydraulic) %>%
  dplyr::filter(length >= 0.05) %>% # ignore tubing lengths < 5cm
  ggplot(aes(x = length*1e2, y = q * 1e9, color = type)) + # convert m^3/s into uL/s
  geom_line() + 
  scale_color_viridis_d(labels = c("PTFE", "Tygon")) + 
  labs(x = "Tubing length [cm]",
       y = "Flow rate at 1 Pa [uL/s]",
       color = "Tubing type") + 
  theme_pretty()
```







## Straight rectangular channel

TODO



## Networks

TODO

## References
