---
title: "Log Files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Log Files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(picoinjector)
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
```


What do I want?

- plot with q_oil, q_emulsion and p_injection on the y-axis, time on the x-axis
- function to give me pressure and flowrates for a given time

Challenge:

- align pressure and syringe pump profiles such that the timesteps are the same



Overall workflow:

1. Read in the log files
2. Find values of elapsed time of both syringe and pressure log in order to align them
3. Plot


### syringe pump

Read in data:


```{r}
syringe <- logfile_read_syringepump("data/logfiles/syringe_pump_log_20220805_150634.csv")
syringe
```


```{r}
logfile_plot_syringepump(syringe)
```

```{r}
logfile_plot_syringepump(syringe) +
  xlim(c(81.5, 82))
```

Time when q2 was set to zero:

```{r}
syringe %>% 
  filter(channel == "q2",
         elapsed_time > 4904 & elapsed_time < 4905)
```







### pressure controller

A drawback of the pressure controller log files is, that the system time is not stored.
Therefore, we need to align the pressure profile with the syringe pump profile in order to get that information.



```{r}
pressure <- logfile_read_pressure("data/logfiles/pressure_log.txt")
pressure
```

```{r}
logfile_plot_pressure(pressure)
```

Find time when p2 was set to zero:

```{r}
pressure %>% 
  filter(channel == "p2",
         value == "target",
         elapsed_time > 4861.5 & elapsed_time < 4862.5)
```



### align profiles

What are the times where pressure and syringe pump logs should be aligned?

```{r}
tPressure <- 4862.1
tSyringe <- 4904.6
```

Align pressure log to syringe pump log:

```{r}
pressureAligned <- logfile_align_pressure(pressure, syringe, tPressure, tSyringe)
pressureAligned
```



### plot

Remove unused channels:

```{r}
syringe <- syringe %>% 
  dplyr::filter(channel != "q3")
pressureAligned <- pressureAligned %>% 
  dplyr::filter(channel == "p2")
```



```{r}
# find out by how much the second y-axis has to be scaled:
r <- max(syringe$flowrate) / max(pressureAligned$mbar)
```

Merge pressure and syringe dataframes:

```{r}
df1 <- syringe %>% 
  dplyr::rename(value = flowrate) %>% 
  dplyr::mutate(type = "syringe pump")

df2 <- pressureAligned %>% 
  dplyr::filter(value == "target") %>% 
  dplyr::select(!value) %>% 
  dplyr::rename(value = mbar) %>% 
  dplyr::mutate(type = "pressure controller")

# scale pressure
df2 <- df2 %>% 
  dplyr::mutate(value = value * r)


df <- dplyr::bind_rows(df1, df2)
```


```{r}
ggplot(df, aes(x = elapsed_time / 60, y  = value, color = channel, linetype = type)) + 
  geom_line() + 
  scale_y_continuous(sec.axis = sec_axis(trans = ~./r, name = "Pressure [mbar]")) + 
  theme_pretty() + 
  labs(y = "Flowrate [ul/h]",
       x = "Elapsed time [min]",
       linetype = "Instrument",
       color = "Channel")
```



